---
title: "Flood Detection Analysis"
author: "The Indo Guys (and Gal)"
date: "August 31, 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(caret)
```

## Background

Jakarta flood has become somekind of a tradition, where flood can happen yearly and can go for over 2 weeks on every occurence
Research found that there are a few reasons for the flood: 
- high precipitation
- humidity level
- tide level
- declining land height

## Problem Statement {.flexbox .vcenter}

Can we predict if flooding is going to happen given the weather data H-1 and H-2 prior of the event itself?

## Expected Result {.flexbox .vcenter}

To confirm if there is statistical significance that we are able predict the flood before it happens, based on the weather data

## Data Sources

https://data.go.id/dataset/curah-hujan-dan-hari-hujan-menurut-bulan-dki-jakarta
https://bpbd.jakarta.go.id/waterlevel/
http://data.jakarta.go.id/group/lingkungan-hidup

## Basic data from NOAA {.flexbox .vcenter}


```{r}
banjir_data <- read.csv("data/CLEAN DATASET BANJIR.csv", stringsAsFactors = FALSE)

banjir_data$PREDH1 <- factor(banjir_data$PREDH1, levels = c("0", "1"), labels = c("Normal", "Banjir"))
banjir_data$PREDH2 <- factor(banjir_data$PREDH2, levels = c("0", "1", "2", "3"), labels = c("Normal", "H-2", "H-1", "Banjir"))

banjir_data$MXSPD <- as.numeric(banjir_data$MXSPD)
banjir_data$PRCP <- as.numeric(banjir_data$PRCP)
```

Now, let's use PREDH1, remove PREDH2
```{r}
banjir_data <- subset(banjir_data, select = -c(STN, YEARMODA, PREDH2, FOG, SNOW, HAIL))

set.seed(3)

banjir_data <- banjir_data[sample(nrow(banjir_data)),]

table(banjir_data$PREDH1)

```



```{r}
# Creating a normalize() function, which takes a vector x and for each value in that vector, subtracts the minimum value in x and divides by the range of x

normalize <- function(x){
  return ( 
    (x - min(x))/(max(x) - min(x)) 
           )
}
```


Normalize data
```{r}
#banjir_data_n <- as.data.frame(lapply(banjir_data[,1:13], normalize))

#banjir_data_x <- merge(banjir_data_n, banjir_data$PREDH1, by="")
# n0_var <- nearZeroVar(banjir_data[,1:14])
# banjir_data <- banjir_data[,-n0_var]

banjir_data_intrain <- sample(nrow(banjir_data), nrow(banjir_data)*0.8)
banjir_data_train <- banjir_data[banjir_data_intrain, ]
banjir_data_test <- banjir_data[-banjir_data_intrain, ]

```


```{r}
prop.table(table(banjir_data_train$PREDH1))

prop.table(table(banjir_data_test$PREDH1))
```


```{r}
ctrl <- trainControl(method="repeatedcv", number=5, repeats=3)
banjir_data_forest <- train(PREDH1 ~ ., data=banjir_data_train, method="rf", trControl = ctrl)
```



```{r}
banjir_data_forest
```

```{r}
table(predict(banjir_data_forest, banjir_data_test[,-14]), banjir_data_test[,14])
```




## Data with Pintu Air (water gate data)

Here are the data sources

```{r}
banjir_data <- read.csv("data/CLEAN DATASET BANJIR2.csv", stringsAsFactors = FALSE)

banjir_data$PREDH1 <- factor(banjir_data$PREDH1, levels = c("0", "1"), labels = c("Normal", "Banjir"))
banjir_data$PREDH2 <- factor(banjir_data$PREDH2, levels = c("0", "1", "2", "3"), labels = c("Normal", "H-2", "H-1", "Banjir"))

banjir_data$MXSPD <- as.numeric(banjir_data$MXSPD)
banjir_data$PRCP <- as.numeric(banjir_data$PRCP)
```

## Cleaning up data

Removing unnecessary data sets, such as Station, Year-Month-Day, Fog, Snow and Hail data
```{r}
banjir_data <- subset(banjir_data, select = -c(STN, YEARMODA, PREDH2, FOG, SNOW, HAIL))
```



```{r}
set.seed(3)

banjir_data <- banjir_data[sample(nrow(banjir_data)),]

table(banjir_data$PREDH1)

```



```{r}
# Creating a normalize() function, which takes a vector x and for each value in that vector, subtracts the minimum value in x and divides by the range of x

normalize <- function(x){
  return ( 
    (x - min(x))/(max(x) - min(x)) 
           )
}
```


Normalize data
```{r}
#banjir_data_n <- as.data.frame(lapply(banjir_data[,1:13], normalize))

#banjir_data_x <- merge(banjir_data_n, banjir_data$PREDH1, by="")
# n0_var <- nearZeroVar(banjir_data[,1:14])
# banjir_data <- banjir_data[,-n0_var]

banjir_data_intrain <- sample(nrow(banjir_data), nrow(banjir_data)*0.8)
banjir_data_train <- banjir_data[banjir_data_intrain, ]
banjir_data_test <- banjir_data[-banjir_data_intrain, ]

```


```{r}
prop.table(table(banjir_data_train$PREDH1))

prop.table(table(banjir_data_test$PREDH1))
```


```{r}
ctrl <- trainControl(method="repeatedcv", number=5, repeats=3)
banjir_data_forest <- train(PREDH1 ~ ., data=banjir_data_train, method="rf", trControl = ctrl)
```



```{r}
banjir_data_forest
```

```{r}
table(predict(banjir_data_forest, banjir_data_test[,-15]), banjir_data_test[,15])
```


