---
title: "Flood Detection Analysis"
author: "The Indo Guys (and Gal)"
date: "August 31, 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
set.seed(1)
```

## Problem



## Data source from NOAA


Data source sample
```{r}
banjir_data <- read.csv("data/CLEAN DATASET BANJIR.csv", stringsAsFactors = FALSE)
head(banjir_data)
```

## Data source explanation
Breakdown of the data structure

CODE | Description
-----|-----:
STN | Station number for the location.
YEARMODA | The year, month and day.
TEMP | Mean temperature for the day in degrees Fahrenheit to tenths.
DEWP | Mean dew point for the day in degrees Fahrenheit to tenths.
SLP | Mean sea level pressure for the day in millibars to tenths.


CODE | Description
-----|-----:
STP | Mean station pressure for the day in millibars to tenths.
VISIB | Mean visibility for the day in miles to tenths.     
WDSP | Mean wind speed for the day in knots to tenths.
MXSPD | Maximum sustained wind speed reported for the day in knots to tenths.
GUST | Maximum wind gust reported for the day in knots to tenths.

CODE | Description
-----|-----:
MAX | Maximum temperature reported during the day in Fahrenheit to tenths--time of max temp report varies by country and region, so this will sometimes not be the max for the calendar day.
MIN | Minimum temperature reported during the day in Fahrenheit to tenths--time of min  temp report varies by country and region, so this will sometimes not be the min for the calendar day. 
PRCP | Total precipitation (rain and/or melted snow) reported during the day in inches and hundredths.
SNDP | Snow depth in inches to tenths--last report for the day if reported more than once.
FRSHTT | Indicators for the occurrence during the day of: Fog ('F' - 1st digit). Rain or Drizzle ('R' - 2nd digit). Snow or Ice Pellets ('S' - 3rd digit). Hail ('H' - 4th digit). Thunder ('T' - 5th digit). Tornado or Funnel Cloud ('T' - 6th digit).



```{r, echo=FALSE}
banjir_data$PREDH1 <- factor(banjir_data$PREDH1, levels = c("0", "1"), labels = c("Normal", "Banjir"))
banjir_data$PREDH2 <- factor(banjir_data$PREDH2, levels = c("0", "1", "2", "3"), labels = c("Normal", "H-2", "H-1", "Banjir"))

banjir_data$MXSPD <- as.numeric(banjir_data$MXSPD)
banjir_data$PRCP <- as.numeric(banjir_data$PRCP)
```

Our source of data consist of banjir (Flood) and Normal (no flooding)
```{r}
table(banjir_data$PREDH1)
```

## Data preparation

We remove unnecessary data
```{r}
banjir_data <- subset(banjir_data, select = -c(STN, YEARMODA, PREDH2, FOG, SNOW, HAIL))
```

Also, we randomize from time series to avoid uneven spread data
```{r}

banjir_data <- banjir_data[sample(nrow(banjir_data)),]

```



```{r, echo = false}
# Creating a normalize() function, which takes a vector x and for each value in that vector, subtracts the minimum value in x and divides by the range of x

normalize <- function(x){
  return ( 
    (x - min(x))/(max(x) - min(x)) 
           )
}
```


We split the data into training data (80%) and test data (20%)
```{r}
#banjir_data_n <- as.data.frame(lapply(banjir_data[,1:13], normalize))

#banjir_data_x <- merge(banjir_data_n, banjir_data$PREDH1, by="")
# n0_var <- nearZeroVar(banjir_data[,1:14])
# banjir_data <- banjir_data[,-n0_var]

banjir_data_intrain <- sample(nrow(banjir_data), nrow(banjir_data)*0.8)
banjir_data_train <- banjir_data[banjir_data_intrain, ]
banjir_data_test <- banjir_data[-banjir_data_intrain, ]

```
## The data ratio between the train data and test data

```{r}
prop.table(table(banjir_data_train$PREDH1))

prop.table(table(banjir_data_test$PREDH1))
```
Using set.seed(1), we see even spread of Normal and Banjir for both of data

## Random forest function

Here we use RepeatedCV
```{r}
ctrl <- trainControl(method="repeatedcv", number=5, repeats=3)
banjir_data_forest <- train(PREDH1 ~ ., data=banjir_data_train, method="rf", trControl = ctrl)
```


## Random forest

Result of the random forest function
```{r}
banjir_data_forest
```


## Prediction based on the formula

We try to see the result of the prediction
```{r}
banjir_rf <- table(predict(banjir_data_forest, banjir_data_test[,-14]), banjir_data_test[,14])

banjir_rf
```


Accuracy is:
```{r}
(banjir_rf["Normal", "Normal"] + banjir_rf["Banjir", "Banjir"]) / sum(banjir_rf)
```

## What if

## Data with Pintu Air (water gate data)

Here are the data sources

```{r}
banjir_data <- read.csv("data/CLEAN DATASET BANJIR2.csv", stringsAsFactors = FALSE)

banjir_data$PREDH1 <- factor(banjir_data$PREDH1, levels = c("0", "1"), labels = c("Normal", "Banjir"))
banjir_data$PREDH2 <- factor(banjir_data$PREDH2, levels = c("0", "1", "2", "3"), labels = c("Normal", "H-2", "H-1", "Banjir"))

banjir_data$MXSPD <- as.numeric(banjir_data$MXSPD)
banjir_data$PRCP <- as.numeric(banjir_data$PRCP)
```

## Cleaning up data

Removing unnecessary data sets, such as Station, Year-Month-Day, Fog, Snow and Hail data
```{r}
banjir_data <- subset(banjir_data, select = -c(STN, YEARMODA, PREDH2, FOG, SNOW, HAIL))
```



```{r}
set.seed(3)

banjir_data <- banjir_data[sample(nrow(banjir_data)),]

table(banjir_data$PREDH1)

```



```{r}
# Creating a normalize() function, which takes a vector x and for each value in that vector, subtracts the minimum value in x and divides by the range of x

normalize <- function(x){
  return ( 
    (x - min(x))/(max(x) - min(x)) 
           )
}
```


Normalize data
```{r}
#banjir_data_n <- as.data.frame(lapply(banjir_data[,1:13], normalize))

#banjir_data_x <- merge(banjir_data_n, banjir_data$PREDH1, by="")
# n0_var <- nearZeroVar(banjir_data[,1:14])
# banjir_data <- banjir_data[,-n0_var]

banjir_data_intrain <- sample(nrow(banjir_data), nrow(banjir_data)*0.8)
banjir_data_train <- banjir_data[banjir_data_intrain, ]
banjir_data_test <- banjir_data[-banjir_data_intrain, ]

```


```{r}
prop.table(table(banjir_data_train$PREDH1))

prop.table(table(banjir_data_test$PREDH1))
```


```{r}
ctrl <- trainControl(method="repeatedcv", number=5, repeats=3)
banjir_data_forest <- train(PREDH1 ~ ., data=banjir_data_train, method="rf", trControl = ctrl)
```



```{r}
banjir_data_forest
```

```{r}
table(predict(banjir_data_forest, banjir_data_test[,-15]), banjir_data_test[,15])
```


